Volunteers in the Clouds: an Architecture for Low Cost and Potentially Massively Distributed Evolutionary Computation

### 2015-07-27 volunteer-clouds-ieee-ci-si.tex
Answering @mariosky and accepting his changes

The web used as a resource for distributed computing has a
long history, and almost since the beginning it was used for 
evolutionary computation even before JavaScript was
introduced as a browser-based scripting language in 1997. Before that
time other procedures,
%other technologies?
 including Flash animations and VBScript, ActiveX or Java applets were
used; for instance, Java was pointed out in \cite{soares1998get} as a ``language for
internet'', providing some advantages such as multi-architecture compatibility or 
security mechanisms and was proposed in \cite{chandy1996world} as the
basis for the creation of a world-wide peer to peer system using Java
compiled programs. The ATLAS system
\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
with another language called Cilk; all these papers indicated that
the introduction of an architecture-independent system such as the
Java virtual machine opened the doors for ultrascale, Internet-wide,
metacomputers.

Using any of these languages allowed the creation
of client-server applications in the browser, provided that the user had
installed the required virtual machine plugin first, but with the advantage of
providing a single and universal way of accessing resources via an
URI, something that is not guaranteed using desktop applications. However, this style of
computing presents several issues and challenges which were addressed
as early as 1996 by Baratloo et al. in their system Charlotte:
\cite{baratloo1996charlotte}. This system, which implemented a
distributed shared memory, was inspired by previous work
on parallel computing systems PVM and MPI, but solved most issues by
using the Java virtual machine embedded in browser,
% by using the browser's embedded Java virtual machine,
which provides
security to the user, uniformity of computing environment
independently of the operating system and  access to a common data store in the
server. Other issues such as scheduling are solved on the server. This
paper was later extended and published in a journal in 1999
\cite{baratloo1999charlotte}, but it is interesting to note that
already in the early stages of the web the possibility of
%already in the early stages of the web, the possibility of
browser-based distributed computing was already researched.

One of the first papers by Soares et al. used Java, and in their paper \cite{soares1998get} they
describe JET, a system that supports 
the execution of parallel applications over the Internet. This system has 
a comprehensive statistics support, allowing its use for science,
since it allows the collection of algorithm analytics by the
experiment designer, unlike the other systems compared in that
%experiment designer, unlike other systems compared in that
paper. Contemporary proposals, such as SuperWeb
\cite{alexandrov1997superweb} also propose a translation of classical
distribution techniques, with complicated architectures that include
brokers for resource discovery. It is focused, however, in solving the
issues involved in browser-based computing and also in making a system
that is able to work across different classes of computers, mentioning
SPARCstations and Windows-95 computers. Interestingly enough, this
paper incorporates a discussion on the economic model, paving the way
for the for-profit distributed computing systems that later on
arose. On the other hand, the potential for volunteer computing using
browsers was realized 
later on \cite{sarmenta-bayanihan} as well as the potential for
mischief of users with code in their hands
\cite{sarmenta-sabotagetolerance}. However, these early efforts by
Sarmenta once again used Java and not JavaScript, making this effort
less universal.

@@

%Si no se extiende mucho se podría también hablar de la conexión con Cloud Computing,
%se discuten las ventajas de utilizar Javascript para acceder a recursos donados por voluntarios
%pero falta hablar de como se explotaría el recurso Cloud. Lo digo por el titulo de Volunteers in the Clouds. Tal vez
%se deba hablar de la posibilidad de escalar los sistemas a bajo costo o según la demanda.

>>>>

The web used as a resource for distributed computing has a
long history, and almost since the beginning it was used for 
evolutionary computation even before JavaScript was
introduced as a browser-based scripting language in 1997. Before that
time other technologies,
%other technologies?
% changed - JJ
 including Flash animations and VBScript, ActiveX or Java applets were
used; for instance, Java was pointed out in \cite{soares1998get} as a ``language for
internet'', providing some advantages such as multi-architecture compatibility or 
security mechanisms and was proposed in \cite{chandy1996world} as the
basis for the creation of a world-wide peer to peer system using Java
compiled programs. The ATLAS system
\cite{Baldeschwieler:1996:TIG:504450.504482} also used Java together
with another language called Cilk; all these papers indicated that
the introduction of an architecture-independent system such as the
Java virtual machine opened the doors for ultrascale, Internet-wide,
metacomputers.

Using any of these languages allowed the creation
of client-server applications in the browser, provided that the user had
installed the required virtual machine plugin first, but with the advantage of
providing a single and universal way of accessing resources via an
URI, something that is not guaranteed using desktop applications. However, this style of
computing presents several issues and challenges which were addressed
as early as 1996 by Baratloo et al. in their system Charlotte:
\cite{baratloo1996charlotte}. This system, which implemented a
distributed shared memory, was inspired by previous work
on parallel computing systems PVM and MPI, but solved most issues by
using the Java virtual machine embedded in the browser,
which provides
security to the user, uniformity of computing environment
independently of the operating system and  access to a common data store in the
server. Other issues such as scheduling are solved on the server. This
paper was later extended and published in a journal in 1999
\cite{baratloo1999charlotte}, but it is interesting to note that
already in the early stages of the web the possibility of
browser-based distributed computing was already researched.

One of the first papers by Soares et al. used Java, and in their paper \cite{soares1998get} they
describe JET, a system that supports 
the execution of parallel applications over the Internet. This system has 
a comprehensive statistics support, allowing its use for science,
since it allows the collection of algorithm analytics by the
experiment designer, unlike the other systems compared in that
paper. Contemporary proposals, such as SuperWeb
\cite{alexandrov1997superweb} also propose a translation of classical
distribution techniques, with complicated architectures that include
brokers for resource discovery. It is focused, however, in solving the
issues involved in browser-based computing and also in making a system
that is able to work across different classes of computers, mentioning
SPARCstations and Windows-95 computers. Interestingly enough, this
paper incorporates a discussion on the economic model, paving the way
for the for-profit distributed computing systems that later on
arose. On the other hand, the potential for volunteer computing using
browsers was realized 
later on \cite{sarmenta-bayanihan} as well as the potential for
mischief of users with code in their hands
\cite{sarmenta-sabotagetolerance}. However, these early efforts by
Sarmenta once again used Java and not JavaScript, making this effort
less universal.

@@

%Si no se extiende mucho se podría también hablar de la conexión con Cloud Computing,
%se discuten las ventajas de utilizar Javascript para acceder a recursos donados por voluntarios
%pero falta hablar de como se explotaría el recurso Cloud. Lo digo por el titulo de Volunteers in the Clouds. Tal vez
%se deba hablar de la posibilidad de escalar los sistemas a bajo costo
%o según la demanda.

% Si te animas, propón un párrafo - JJ
